### 2025-08-07 README.Rmd
clean ups for release

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "ragg_png",
  dpi = 96
)
options(tibble.print_min = 5, tibble.print_max = 5)
library("mgcv")
library("gratia")
```

# gratia

<!-- badges: start -->
[![R build status](https://github.com/gavinsimpson/gratia/workflows/R-CMD-check/badge.svg)](https://github.com/gavinsimpson/gratia/actions)
[![codecov](https://codecov.io/gh/gavinsimpson/gratia/branch/main/graph/badge.svg?token=GG5NQfgRFu)](https://app.codecov.io/gh/gavinsimpson/gratia)
[![CRAN\_Status\_Badge](https://www.r-pkg.org/badges/version/gratia)](https://cran.r-project.org/package=gratia)
[![CRAN Downloads](https://cranlogs.r-pkg.org/badges/grand-total/gratia)](https://cran.r-project.org/package=gratia)
[![DOI](https://joss.theoj.org/papers/10.21105/joss.06962/status.svg)](https://doi.org/10.21105/joss.06962)
[![r-universe version](https://gavinsimpson.r-universe.dev/gratia/badges/version)](https://gavinsimpson.r-universe.dev/gratia)
[![r-universe status](https://gavinsimpson.r-universe.dev/gratia/badges/checks)](https://gavinsimpson.r-universe.dev/gratia)
<!-- badges: end -->

## Overview

Working with GAMs within the 'tidyverse' can be tedious and even difficult
without a good understanding of GAMs themselves and how the model is returned
by 'mgcv' and what the model objects contain. 'gratia' is designed to help with
this.

'gratia' provides 'ggplot'-based graphics and utility functions for working with
generalized additive models (GAMs) fitted using the 'mgcv' package, via a
reimplementation of the `plot()` method for GAMs that 'mgcv' provides, as well
as 'tidyverse' compatible representations of estimated smooths.

## Citation

If you use *gratia* in your work, please cite the package using:

Simpson, G. L., (2024). gratia: An R package for exploring generalized additive
models. *Journal of Open Source Software*, **9**(104), 6962,
<https://doi.org/10.21105/joss.06962>

## Features

```{r draw-gam-figure, fig.height = 6, fig.width = 9, echo = FALSE, fig.show = "hide", message = FALSE}
df1 <- data_sim("eg1", n = 400, seed = 42)
m1 <- gam(y ~ s(x0) + s(x1) + s(x2) + s(x3), data = df1, method = "REML")
draw(m1, residuals = TRUE, rug = FALSE)
```

```{r draw-gam-figure-2d, fig.height = 6, fig.width = 9, echo = FALSE, fig.show = "hide", message = FALSE}
df2 <- data_sim("eg2", n = 1000, seed = 42)
m2 <- gam(y ~ s(x, z), data = df2, method = "REML")
draw(m2)
```

```{r appraise-figure, fig.height = 6, fig.width = 9, echo = FALSE, fig.show = "hide"}
appraise(m1)
```

The main features of *gratia* are currently

* A *ggplot2*-based replacement for `mgcv:::plot.gam()`: `draw.gam()`.

  For example, the classic four term additive example from Gu & Wahba:

  ![Estimated smooths from a GAM](man/figures/README-draw-gam-figure-1.png)

  Or for a bivariate smooth:

  ![Estimated smooths from a GAM](man/figures/README-draw-gam-figure-2d-1.png)

  Note that some specialist smoothers (`bs %in% c("mrf","sw", "sf")`) are not
  currently supported, but univariate, *factor* and *continuous* `by`-variable
  smooths, simple random effect smooths (`bs = 're'`), factor-smooth
  interaction smooths (`bs = "fs"`), constrained factor smooths (`bs = "sz"`),
  full soap film smooths (`bs = "so"`), and bivariate, trivariate, and
  quadvariate TPRS and tensor product smooths are supported,

* Estimation of derivatives of fitted smoothers: `derivatives()`,

* Estimation of point-wise across-the-function confidence intervals and
  simultaneous intervals for smooths: `confint.gam()`.

* Model diagnostics via `appraise()`

  ![Model diagnostics figure](man/figures/README-appraise-figure-1.png)

## Installing *gratia*

*gratia* is now available on CRAN, and can be installed with

```{r install, eval = FALSE}
install.packages("gratia")
```

however *gratia* is under active development and you may wish to install the
development version from github. The easiest way to do this is via the
`install_github()` function from package *remotes*. Make sure you have
*remotes* installed, then run

```{r install-github, eval = FALSE}
remotes::install_github("gavinsimpson/gratia")
```

to install the package. Alternatively, binary packages of the development
version are available from rOpenSci's R Universe service:

```{r install-r-universe, eval = FALSE}
# Install gratia in R
install.packages("gratia", repos = c(
  "https://gavinsimpson.r-universe.dev",
  "https://cloud.r-project.org"
))
```

## History

*gratia* grew out of an earlier package, *schoenberg*, itself a development of
the earlier package *tsgam*, which was originally intended to be used with GAMs
fitted to time series. As I was developing *tsgam* however it became clear that
the package could be used more generally and that the name "tsgam" was no longer
appropriate. To avoid breaking blog posts I had written using *tsgam* I decided
to copy the git repo and all the history to a new repo for the package under the
name *schoenberg*. At a later date someone released another package called
*schoenberg* to CRAN, so that scuppered that idea. Now I'm calling the package
*gratia*. Hopefully I won't have to change it again&hellip;

## Why *gratia*?

In naming his [*greta*](https://github.com/greta-dev/greta) package, Nick
Golding observed the recent phenomena of naming statistical modelling software,
such as Stan or Edward, after individuals that played a prominent role in the
development of the field. This lead Nick to name his Tensor Flow-based package
*greta* after [*Grete Hermann*](https://greta-stats.org/articles/webpages/why_greta.html).

In the same spirit, *gratia* is named in recognition of the contributions of
[Grace Wahba](https://en.wikipedia.org/wiki/Grace_Wahba), who did pioneering
work on the penalised spline models that are at the foundation of the way GAMs
are estimated in *mgcv*. I wanted to name the package *grace*, to explicitly
recognise Grace's contributions, but unfortunately there was already a package
named *Grace* on CRAN. So I looked elsewhere for inspiration.

The English word "grace" derives from the Latin *gratia*, meaning "favor, charm,
thanks" ([according to Merriam Webster](https://www.merriam-webster.com/dictionary/grace)).

The chair that Grace Wabha currently holds is named after 
[Isaac J Schoenberg](https://en.wikipedia.org/wiki/Isaac_Jacob_Schoenberg), a
former University Madison-Wisconsin Professor of Mathematics, who in a 1946
paper provided the first mathematical reference to "splines". (Hence the
previous name for the package.)